Boosted Mirror Descent, Frank-Wolfe, and Herding

### 2013-10-04 icml/lower-bounds.tex
more small fixes; tried to improve l.b. section

\begin{document}
\section{Lower Bounds on Convergence Rates}
\label{sec:lower-bounds}

It has been an open question whether it is possible to achieve the fast $O(1/T^{2})$ convergence of finite-dimensional herding in the infinite-dimensional case. We now show that this is not in general possible. In particular, the upper bounds in the previous section for herding and herding-like algorithms have matching lower bounds for at least some combinations of distributions and feature functions.

For the matching $O(1/T)$ lower bound we analyze sparse update methods. Formally, define a {\em sparse update method} to be any method that produces an estimator of $\bar \phi$ at iteration $t$ of the form 
\[
\theta_{t} = \sum_{s =1}^{t}w_{t} \phi(x_{t}).
\]
Sparse update methods can be viewed as characterizing ``herding-like'' algorithms, including the generalized herding and \cgd with line search. These sparse methods -- which only operate on the extreme points of the simplex -- are of interest because they provide (weighted) ``samples'' matching the moments of the distribution of interest.

The bound setup is as follows. Let $\sX$ be the set of vectors $(x_{1},x_{2},\dots) \in [0,1]^{\infty}$ such that $\sum_{n=1}^{\infty} x_{n} = 1$ and a finite number of the $x_{n}$ are non-zero. Define the infinite feature vector $\phi$ to be the imbedding of $\sX$ into $\ell^{1}$. Let the target distribution $p(x)$ be chosen such that $\bar\phi_{n} = c/\alpha(n)$, where $\alpha(n)$ is any superlinear function and $c^{-1} = \sum_{n=1}^{\infty} \alpha(n) < \infty$. Without loss of generality we can assume that $\alpha$ is monotonically increasing. We will consider the standard $\ell^{2}$ herding loss function $L(\theta) = \|\theta - \bar \phi\|_2^2$.

>>>>

\begin{document}
\section{A General Lower Bound on Convergence Rates}
\label{sec:lower-bounds}

Although it is possible to achieve the fast $O(1/T^{2})$ convergence rate for finite-dimensional herding, \citet{Bach:2012a} note that without further assumptions herding has only a $O(1/T)$ convergence rate in the infinite-dimensional case. We now provide an example showing rate is tight for {\em any} herding-like algorithm.  By ``herding-like,'' we mean algorithms that approximate the distribution with at most one additional sample at each step. We formalize this idea by defining a {\em sparse update method} to be any method that produces an estimator of $\bar \phi$ at iteration $t$ of the form 
\[
\theta_{t} = \sum_{s =1}^{t}w_{t} \phi(x_{t}).
\]
Sparse update methods include $\alpha_{t}$-weighted herding and \cgd with line search. These sparse methods -- which only operate on the extreme points of the simplex -- are of interest because they provide (weighted) ``samples'' matching the moments of the distribution of interest.

Our example is as follows. Let $\sX$ be the set of vectors $(x_{1},x_{2},\dots) \in [0,1]^{\infty}$ such that $\sum_{n=1}^{\infty} x_{n} = 1$ and a finite number of the $x_{n}$ are non-zero. Define the infinite feature vector $\phi$ to be the imbedding of $\sX$ into $\ell^{1}$. Let the target distribution $p(x)$ be chosen such that $\bar\phi_{n} = c/\alpha(n)$, where $\alpha(n)$ is any superlinear function and $c^{-1} = \sum_{n=1}^{\infty} \alpha(n) < \infty$. Without loss of generality we can assume that $\alpha$ is monotonically increasing. We will consider the standard $\ell^{2}$ herding loss function $L(\theta) = \|\theta - \bar \phi\|_2^2$.

### 2013-10-04 icml/paper.tex
first draft of introduction
intro improved; log(t)/t writeup added
added bib filea
split of bmd section. added shortcuts for algorithm names
draft of connections subsection
added smc and mcmc to bibliography
more small fixes; tried to improve l.b. section

\documentclass{article}
\usepackage{import, subfiles}
\usepackage{hyperref, icml2014}
\input{latex-defs.tex}

@@

% You may provide any keywords that you 
% find helpful for describing your paper; these are used to populate 
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{boosting, optimization, mirror descent, frank-wolfe, herding}

@@

\begin{abstract} 
TODO
\end{abstract}

\subfile{introduction.tex}
\subfile{herding-connection.tex}
\subfile{convergence-proofs.tex}
%\subfile{lower-bounds.tex}
%\subfile{chen-proofs.tex}
%\subfile{conclusion.tex}

%\bibliography{example_paper}
%\bibliographystyle{icml2013}

>>>>

\documentclass{article}
\usepackage{import, subfiles}
\usepackage{hyperref, icml2014}
\usepackage{tabularx}
\input{latex-defs.tex}

@@

% You may provide any keywords that you 
% find helpful for describing your paper; these are used to populate 
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{boosting, optimization, mirror descent, frank-wolfe, conjugate gradient descent, herding}

@@

\begin{abstract} 
We introduce a unifying framework for understanding herding and 
related algorithms. The framework is based on pair of optimization 
algorithms we call primal and dual boosted mirror descent.
We define two classes of generalized herding algorithms, 
including one for infinite dimensional spaces
that may apply even when herding does not. A number 
of convergence bounds that apply to herding and generalized
herding are obtained. Along the way we provide
a more cohesive picture of the connections between herding, 
conditional gradient descent and related methods, and maximum entropy. 
\end{abstract}

\subfile{introduction.tex}
\subfile{boosted-mirror-descent.tex}
\subfile{herding-connection.tex}
\subfile{convergence-proofs.tex}
\subfile{lower-bounds.tex}
%\subfile{chen-proofs.tex}
\subfile{conclusion.tex}

\bibliography{herding.bib}
\bibliographystyle{icml2014}
